---
title: "Introduction"
author: "Colin O'Rourke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{lme4, magrittr}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Set a seed for reproducibility.

```{r set-a-seed}
set.seed(1313L)
```

Activate useful libraries.

```{r libs}
library(lme4)
library(magrittr)
library(npicc)
```

# Purpose

Here we want to introduce the calculation of a measure of group association called the Intraclass Concordance Coefficient. This quantity is similar in flavour to the commonly used Intraclass Correlation Coefficient, and may be view as its nonparametric, rank-based analogue.

# Simulate Data

To demonstrate, we will first simulate some data with a known level of association. Here this is done using the `makeSample` function, which simulates data using a gaussian-based components-of-variance structure. In other words the outcome value $j$ for subject $i$, we have,

$$
Y_{ij} = \mu + \eta_{i} + \epsilon_{ij},
$$
where $\eta_i \sim N(0,\tau^2)$ and $\epsilon_{ij} \sim N(0,\sigma^2)$, with $\eta_i \perp \epsilon_{ij}$. Here, $\tau$ is the subject-level SD, and $\sigma$ is the observation-level SD.

Now, we can simulate data from this model, with an overall mean of about -2.3, subject-level SD of 1, and observation-level SD of 0.25. We also want to treat this data as latent, which the actually observed outcomes being the exponentiated latent value.

```{r simulate-data}
df = make_sample(
  nSubj = 10,
  nObsPerSubj = c(2,4,6),
  overall_mean = log(0.1),
  subj_sd = 1,
  obs_sd = 0.25,
  obs_trans = exp
)

# turn subj into a factor, levels ordered by mean level
df$subj = reorder(df$subj, df$obs, mean)

#print result
head(df, n = 10L)
```

These are what our simulated observed values look like:

```{r visualise-data, fig.cap="Plot showing observed value of outcome by subject."}
par(mar = c(4.5, 4.5, 1.5, 1.5))

plot(
  obs_trans ~ jitter(as.integer(subj), factor = 0.1),
  data = df,
  pch=16,
  cex=0.2,
  xlab="Subject #",
  ylab="Value of outcome"
)
```

# Calculate ICCs

In addition to showing how one might calculate the new ICC agreement statistic, we might also want to compare its value with the common ICC using `lme4::lmer`.

```{r calculate-icc-values}
# Standard variance components agreement stat
lmer_icc = if(require("lme4") && require("magrittr")) {
  lmer(obs_trans ~ (1|subj), data = df) %>%
  VarCorr %>%
  as.data.frame %>%
  extract2("vcov") %>%
  {.[1] / sum(.)}
} else {
  NA_real_
}

# New concordance-style agreement stat
npicc_icc = getICC(x = df$subj, y = df$obs_trans)
```

The value of the traditional ICC is `r round(lmer_icc, 2)`, whereas the value of the rank-based ICC is `r round(npicc_icc,2)`.
